---
title: "R Notebook"
author: Emre Erhan
date: '2018-11-09'
theme: yeti
output: github_document
---

## Import required packages

```{r results='hide', message=FALSE, warning=FALSE}
library(stringr)
library(gapminder)
library(ggplot2)
library(dplyr)
library(testthat)
```



## 2. Writing a function

In computational biology, when working with DNA sequences, it is often useful to get the "canonical" version of a string of DNA bases. The canonical version is either the lexographically maximal or lexographically minimal complement of the sequence. I will write a function that returns the canonical sequence for a given DNA sequence.

```{r}
canonical = function(seq){
  complement = chartr("ATGC", "TACG", seq) # chartr determines the mapping of complementary basepairs
  if (seq > complement){
    return(seq)
  }
  else{
    return(complement)
  }
}
```

Let's give it a shot! The canonical sequence of "ACGT" and "TGCA" should both be the same, since they're complementary sequences. Let's use `test_that`

```{r}
test_that("A simple test", {
  expect_equal(canonical("ACGT"), canonical("TGCA"))
})
```

It works! Although, it may be useful to specify whether the maximal or minimal complement is required. I'll provide that option in a parameter.

```{r}
canonical = function(seq, minimal = TRUE){
  complement = chartr("ATGC", "TACG", seq) # chartr determines the mapping of complementary basepairs
  if (minimal){
    if (seq < complement){
      return(seq)
    }
    else{
      return(complement)
    }
  }
  else{
    if (seq < complement){
      return(complement)
    }
    else{
      return(seq)
    }
  }
}
```

Let's make some test cases using `test_that`

```{r}
long_seq = 'ACGTCGATCGATGCTAGCTAGCTAGTCGACTATCAGTAG'
test_that("Testing complementary sequences", {
  expect_equal(canonical("ACGT"), canonical("TGCA"))
  expect_equal(canonical(long_seq), canonical(chartr("ATGC", "TACG", long_seq)))
  expect_equal(canonical(long_seq, minimal = FALSE), canonical(chartr("ATGC", "TACG", long_seq), minimal = FALSE))
})
```

Finally, let's make sure that the input is a string, and a DNA sequence.

```{r}
canonical = function(seq, minimal = TRUE){
  if (!is.character(seq)){
    stop(paste("Expecting a string. Was given", typeof(seq)))
  }
  seq = toupper(seq)
  if (!str_detect(seq, '^[ACGT]*$')){
    stop(paste("Expecting a DNA sequence consisting of only A, C, G, T"))
  }
  
  complement = chartr("ATGC", "TACG", seq) # chartr determines the mapping of complementary basepairs
  if (minimal){
    if (seq < complement){
      return(seq)
    }
    else{
      return(complement)
    }
  }
  else{
    if (seq < complement){
      return(complement)
    }
    else{
      return(seq)
    }
  }
}
```

Let's test the error messages.

```{r error=TRUE}
canonical(123)
```

```{r error=TRUE}
canonical("Testing a non-DNA sequence")
```

Excellent! We now have a `canonical` function.